### 2015-07-12 thesis/thesis/abstract.tex
changes

In the last few years, communities of services have been studied in a certain numbers of proposals as virtual pockets of similar expertise. The motivation is to provide these services with high chance of discovery through better visibility, and to enhance their capabilities when it comes to provide requested functionalities. There are a number of proposed mechanisms and models on aggregating web services and making them cooperate within their communities. However, forming optimal and stable communities as coalitions to maximize individual and group efficiency and income for all the involved parties has not been addressed yet. Also, in the proposed frameworks of these communities, a common assumption is that residing services, which are supposed to be autonomous and intelligent, are competing over received requests. However, those services can also exhibit cooperative behaviors, for instance in terms of substituting each other. Deciding to compete or cooperate inside communities is a problem yet to be investigated.

In this thesis, we identify the problem of defining efficient algorithms for coalition formation mechanisms within communities and propose some results using cooperative game-theoretic techniques. We propose a mechanism for community membership requests and selections of web services in the scenarios where there is interaction between one community and many web services and scenarios where web services can join multiple established communities. The ultimate objective is to develop a mechanism for web services to form stable groups allowing them to maximize their efficiency and generate near-optimal (welfare-maximizing) communities. The theoretical and extensive simulation results show that our algorithms provide web services and community owners, in real-world like environments, with applicable and near-optimal decision making mechanisms. We also propose a decision mechanism based on learning methods for services within communities to effectively choose the tasks to perform based on their capabilities and the competition between other community members.
%The experimental results demonstrate the efficiency and scalability of the proposed reduction techniques and show that our work is effective for addressing probabilistic social commitment in MASs.
\\
%\textbf{keywords:}
%Multi-Agent Systems, Agent Communication, Social Commitments, uncertainty, Model Checking.

>>>>

In the last few years, communities of services have been studied in a certain numbers of proposals as virtual pockets of similar expertise. The motivation is to provide these services with high chance of discovery through better visibility, and to enhance their capabilities when it comes to provide requested functionalities. There are a number of proposed mechanisms and models on aggregating web services and making them cooperate within their communities. However, forming optimal and stable communities as coalitions to maximize individual and group efficiency and income for all the involved parties has not been addressed yet. Also, in the proposed frameworks of these communities, a common assumption is that residing services, which are supposed to be autonomous and intelligent, are competing over received requests. However, those services can also exhibit cooperative behaviors, for instance in terms of substituting each other. When competitive and cooperative behaviors and strategies are combined, autonomous services are said to be ``coopetitive''. Deciding to compete or cooperate inside communities is a problem yet to be investigated.

In this thesis, we first identify the problem of defining efficient algorithms for coalition formation mechanisms within communities and propose some results using cooperative game-theoretic techniques. We propose a mechanism for community membership requests and selections of web services in the scenarios where there is interaction between one community and many web services and scenarios where web services can join multiple established communities. Then in order to address the coopetitive relation within our web services, we propose a decision making mechanism for our web services to efficiently choose competition or cooperation strategies to maximize their payoffs. We prove that the proposed decision mechanism is
efficient and can be implemented in time linear in the length of the time period considered for the analysis and the number of services in the community. Moreover, we conduct extensive
simulations, analyze various scenarios, and confirm the obtained theoretical results using parameters from a real web services dataset.

%.  and propose some results using cooperative game-theoretic techniques. We propose a mechanism for community membership requests and selections of web services in the scenarios where there is interaction between one community and many web services and scenarios where web services can join multiple established communities. The ultimate objective is to develop a mechanism for web services to form stable groups allowing them to maximize their efficiency and generate near-optimal (welfare-maximizing) communities. The theoretical and extensive simulation results show that our algorithms provide web services and community owners, in real-world like environments, with applicable and near-optimal decision making mechanisms. We also propose a decision mechanism based on learning methods for services within communities to effectively choose the tasks to perform based on their capabilities and the competition between other community members.
%The experimental results demonstrate the efficiency and scalability of the proposed reduction techniques and show that our work is effective for addressing probabilistic social commitment in MASs.

%\textbf{keywords:}
%Multi-Agent Systems, Agent Communication, Social Commitments, uncertainty, Model Checking.

### 2015-07-12 thesis/thesis/chap1/chap1.tex
changes

Web service communities are dynamic by design
\cite{DBLP:journals/ijebr/MaamarSTBB09}. In these communities, web
services are modeled as intelligent autonomous agents, where they
can adopt a strategy maximizing their payoff at any time. A web
service can ask joining a community and has the right to leave it.
Community managers can invite or ask a web service to leave in
order to maximize the community profit. Users can simply stop
sending requests to a web service which is not providing
satisfactory services. Thus, it is important to consider all the
parties involved in the decision making process about the
community formation and management. Most of the recent work on communities of
services are either user-centric and focus on user satisfaction
\cite{Chun02user-centricperformance} or system-centric and focus
on the whole system throughput, performance and utilization. There
are many contributions in distributed, grid, cluster and cloud
services which are system-centric. However, in real world
environments and applications, both users and service providers
are self-interested agents, aiming to maximize their own profit.
In those environments, both parties (users and services) will
collaborate as long as they are getting more benefits and payoff.
Our initial research question is: \emph{How can we model the
community of agent-based services in order to maximize the utility
of involved users, web services and community organizers?[R1]}

In order to address this problem, recently
\cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed
mechanisms to help users and services maximize their gain. A
two-player non-cooperative game between web services and community master (i.e., manager) was introduced in
\cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In
\cite{DBLP:conf/IEEEscc/LimTMB12}, a 3-way satisfaction approach
for selecting web services has been proposed. In this approach,
the authors proposed a web service selection process that the
community masters can use. The approach considers the efficiency
of all the three involved parties, namely users, web services and
communities. The issue with these solution concepts is that they
consider community as a whole and model it as one entity in their
formulations. A community master decides on behalf of all the
members using an aggregated function of parameters. This can hurt
the overall revenue for some individual web services, or even a
subset of web services. Those services can collude and form their
own community to increase their payoff, instead of having to adjust and share
their resources with other members. Another important issue which
needs to be considered is the community stability. In community of
web services, the members and community organizers collaborate to
perform tasks. Having jointly completed a task and generated
revenue, they need to agree on some reasonable method of dividing
profits (or tasks) among themselves. This is a key issue for the
group stability still to be investigated. If the revenue sharing
mechanism is not fair enough for any subset of web services
working in the community, these agents, as profit maximizing
entities, would deviate and make their own group. Thus, an important an
important research question that we would like to address is:
\emph{How can we model fair and stable communities as coalitions
of agent-based web services?[R2]}

@@

In \cite{10.1109/TSC.2012.12}, a cooperative scheme among
autonomous web services based on coalitional game theory has been
introduced. The authors have proposed an algorithm to reach
individually stable coalition partition for web services in order
to maximize their efficiency. The communities choose new web
services on the promise that it would benefit the community
without decreasing any other web service's income. In their model,
the worth of community is evaluated with high emphasis on the
availability metric and considering price and cost values only.
The community structure is based on a coordination chain, where a
web service is assigned as a \emph{primary} web service and the
community task distribution method will initially invoke the
primary web service. Only if the primary web service is
unavailable, the next backup web services in the ordered
coordination chain will be invoked. However, in cooperative
models, it is preferred to have a real and active cooperative
activity engaging all agents to perform the tasks more
efficiently. Thus, the final research question we aim to investigate is: \emph{How can we model and analyze the cooperation
among the community members in realistic, applicable and practical
settings?[R3]}

Within communities, services can exhibit competitive behavior as they provide the same functionalities and the number of users requests is finite.
However, for the same reason of being functionally similar, services can cooperate with each other, for example to substitute each other in order to perform some sub-tasks.
For instance, services can opt for performing tasks if they feel they are capable enough
or decide to cooperate by showing the availability to perform some sub-tasks. The relevant question to be addressed is: \emph{How can we design a community model where both competitive and cooperative behaviors exist?[R4]}

In most of the recent work on communities of web service, the solutions consider the architecture of centralized management for communities where most of the decisions are made by the centralized coordinator. However, in real world scenarios, decisions made by independent service providers are highly distributed.  \emph{How can we model a distributed decision making process for the problem of forming communities of services?[R5]}

Also in all of the mentioned proposals, the community manager as a centralized entity, has complete information of all the web services and their quality. However, centrality and complete information are strong assumptions, which are not fully compatible with real business scenarios.
\emph{How can we train web services based on limited information to operate efficiently[R6]?}

@@

\section{Research Objectives and Contributions}\label{sec:motexample}

\noindent \emph{\textbf{Contribution \#1:} Efficient Coalition Formation for Web Services\cite{10.1109/TSC.2014.2312940}.}

In this research work, our first objective is to propose a
cooperative model as game for the aggregation of web services
within communities. The solution concepts of our cooperative game
seeks to find efficient ways of forming coalitions (teams) of web
services so that they can maximize their gain and payoff, and
distribute the gain in a fair way among all the member services.
Achieving fairness when the gain is distributed among the
community members is the main factor to keep the coalition stable
as no web service will expect to gain better by deviating from the
community. In other words, the coalition is made efficient if all
the members are satisfied. We first propose a representation
function for communities of web services based on their QoS
attributes. By using this function, we can evaluate the $worth$ of
each community of web services. When facing new membership
requests, a typical community master checks whether the new
coalition having the old and new set of web services will keep the
community stable or not. The community master will reject the
membership requests if it finds out that the new coalition would
be unstable, preventing $any$ subset of web services from gaining
significantly more by deviating from the community and joining
other communities or forming new ones. The computation of
solutions for cooperative games is combinatorial in nature and
proven to be NP-complete \cite{Algorithmic}, making this
computation impractical in scalable real world applications. However, using
the concepts of coalition stability, the second objective is to
investigate approximation algorithms running in polynomial time
providing web services and community masters with applicable and
near-optimal decision making mechanisms.

\noindent \emph{\textbf{Contribution \#2:} Distributed decision making for dynamic formation of web services communities.}

In order to address the centralized decision making process limitations and adopt a distributed decision making approach, we introduce DDM, a Distributed Decision Making model for community formation. DDM regulates web service agents' decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. Unlike existing work on community formation, our decision model is extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and and helps them select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree, which is initially created from the past data, reflects a comprehensive vision about agents' attitudes in terms of their action selection based on their past experiences. Moreover, the tree is getting continuously updated based on both new received feedback and the outcome of chosen actions. This continuous update makes the approach adaptable to any change in the environment.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems.
The decision model provides web services with enough information which helps those services efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in their decision making and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. The proposed model can be seen as a recommneder system that suggests beneficial actions for both communities and single services. Communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact both their short-term and long-term goals; therefore, opting for the best decision available.

@@

\noindent \emph{\textbf{Contribution \#3:} Analyzing Coopetition Strategies of Services within Communities\cite{DBLP:journals/eswa/AslBMKO14}.}

In previous contributions, our focus was on community formation and we emphasised on cooperative behavior of the web services as agents. In our next contribution our focus is on the internal management affairs of communities of web services. Within communities, the web services, selfish and utility maximizers by nature, can follow two different strategies, namely cooperation and competition in order to increase their payoffs when they provide services to consumers \cite{VuFind-10008938119}. In typical business settings, services are used to compete within communities as they provide the same functionalities and the number of users requests is finite. However, the same reason of providing similar functionalities can lead services to cooperate because they can replace each other in case of failure or unavailability, and services can do better in a coalition structure. Analyzing services competition and cooperation strategies within communities is still an open problem that motivates the research described in this section. we propose a mechanism within which service agents
in the community could choose either to compete for an announced task\footnote{Requests and tasks are used in this thesis interchangeably.}, or to cooperate with other competing services in the same community to accomplish some subtasks of the announced task. We equip intelligent web services to follow a reasoning technique to choose best interactive strategy (Coopetitive attitude, which is categorized to compete and cooperate). In the proposed system, We explore details behind the strategic decision making procedures and enable service agents to apply different techniques to constrain high efficiency and obtain the maximum utility. We investigate services' expected payoffs and the involved probabilities that are used to choose over the two interacting strategies.

>>>>

Web service communities are dynamic by design
\cite{DBLP:journals/ijebr/MaamarSTBB09}. In these communities, web
services are modeled as intelligent autonomous agents, where they
can adopt a strategy maximizing their payoff at any time. A web
service can ask joining a community and has the right to leave it.
Community managers can invite or ask a web service to leave in
order to maximize the community profit. Users can simply stop
sending requests to a web service which is not providing
satisfactory services. Thus, it is important to consider all the
parties involved in the decision making process about the
community formation and management. Most of the recent work on communities of
services are either user-centric and focus on user satisfaction
\cite{Chun02user-centricperformance} or system-centric and focus
on the whole system throughput, performance and utilization. There
are many contributions in distributed, grid, cluster and cloud
services which are system-centric. However, in real world
environments and applications, both users and service providers
are self-interested agents, aiming to maximize their own profit.
In those environments, both parties (users and services) will
collaborate as long as they are getting more benefits and payoff.
Our initial research question is: \emph{How can we model the
community of agent-based services in order to maximize the utility
of involved users, web services and community organizers? [R1]}

In order to address this problem, recently
\cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed
mechanisms to help users and services maximize their gain. A
two-player non-cooperative game between web services and community master (i.e., manager) was introduced in
\cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In
\cite{DBLP:conf/IEEEscc/LimTMB12}, a 3-way satisfaction approach
for selecting web services has been proposed. In this approach,
the authors proposed a web service selection process that the
community masters can use. The approach considers the efficiency
of all the three involved parties, namely users, web services and
communities. The issue with these solution concepts is that they
consider community as a whole and model it as one entity in their
formulations. A community master decides on behalf of all the
members using an aggregated function of parameters. This can hurt
the overall revenue for some individual web services, or even a
subset of web services. Those services can collude and form their
own community to increase their payoff, instead of having to adjust and share
their resources with other members. Another important issue which
needs to be considered is the community stability. In community of
web services, the members and community organizers collaborate to
perform tasks. Having jointly completed a task and generated
revenue, they need to agree on some reasonable method of dividing
profits (or tasks) among themselves. This is a key issue for the
group stability still to be investigated. If the revenue sharing
mechanism is not fair enough for any subset of web services
working in the community, these agents, as profit maximizing
entities, would deviate and make their own group. Thus, an important an
important research question that we would like to address is:
\emph{How can we model fair and stable communities as coalitions
of agent-based web services? [R2]}

@@

In \cite{10.1109/TSC.2012.12}, a cooperative scheme among
autonomous web services based on coalitional game theory has been
introduced. The authors have proposed an algorithm to reach
individually stable coalition partition for web services in order
to maximize their efficiency. The communities choose new web
services on the promise that it would benefit the community
without decreasing any other web service's income. In their model,
the worth of community is evaluated with high emphasis on the
availability metric and considering price and cost values only.
The community structure is based on a coordination chain, where a
web service is assigned as a \emph{primary} web service and the
community task distribution method will initially invoke the
primary web service. Only if the primary web service is
unavailable, the next backup web services in the ordered
coordination chain will be invoked. However, in cooperative
models, it is preferred to have a real and active cooperative
activity engaging all agents to perform the tasks more
efficiently. Thus, the final research question we aim to investigate is: \emph{How can we model and analyze the cooperation
among the community members in realistic, applicable and practical
settings? [R3]}

Within communities, services can exhibit competitive behavior as they provide the same functionalities and the number of users requests is finite.
However, for the same reason of being functionally similar, services can cooperate with each other, for example to substitute each other in order to perform some sub-tasks.
For instance, services can opt for performing tasks if they feel they are capable enough
or decide to cooperate by showing the availability to perform some sub-tasks. The relevant question to be addressed is: \emph{How can we design a community model where both competitive and cooperative behaviors exist? [R4]}

In most of the recent work on communities of web service, the solutions consider the architecture of centralized management for communities where most of the decisions are made by the centralized coordinator. However, in real world scenarios, decisions made by independent service providers are highly distributed.  \emph{How can we model a distributed decision making process for the problem of forming communities of services? [R5]}

Also in all of the mentioned proposals, the community manager as a centralized entity, has complete information of all the web services and their quality. However, centrality and complete information are strong assumptions, which are not fully compatible with real business scenarios.
\emph{How can we train web services based on limited information to operate efficiently? [R6]}

@@

\section{Contributions}\label{sec:motexample}

\subsection{Contribution 1: Efficient Coalition Formation for Web Services}

%\cite{10.1109/TSC.2014.2312940}.}

In this research work\footnote{This contribution was published in \cite{10.1109/TSC.2014.2312940}}, our first objective is to propose a
cooperative model as game for the aggregation of web services
within communities. The solution concepts of our cooperative game
seeks to find efficient ways of forming coalitions (teams) of web
services so that they can maximize their gain and payoff, and
distribute the gain in a fair way among all the member services.
Achieving fairness when the gain is distributed among the
community members is the main factor to keep the coalition stable
as no web service will expect to gain better by deviating from the
community. In other words, the coalition is made efficient if all
the members are satisfied. We first propose a representation
function for communities of web services based on their QoS
attributes. By using this function, we can evaluate the $worth$ of
each community of web services. When facing new membership
requests, a typical community master checks whether the new
coalition having the old and new set of web services will keep the
community stable or not. The community master will reject the
membership requests if it finds out that the new coalition would
be unstable, preventing $any$ subset of web services from gaining
significantly more by deviating from the community and joining
other communities or forming new ones. The computation of
solutions for cooperative games is combinatorial in nature and
proven to be NP-complete \cite{Algorithmic}, making this
computation impractical in scalable real world applications. However, using
the concepts of coalition stability, the second objective is to
investigate approximation algorithms running in polynomial time
providing web services and community masters with applicable and
near-optimal decision making mechanisms.

\subsection{Contribution 2: Distributed decision making for dynamic formation of web services communities.}

In order to address the centralized decision making process limitations and adopt a distributed decision making approach, we introduce DDM\footnote{This contribution is submitted to International Journal of Decision Support Systems as a journal paper}, a Distributed Decision Making model for community formation. DDM regulates web service agents' decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. Unlike existing work on community formation, our decision model is extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and and helps them select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree, which is initially created from the past data, reflects a comprehensive vision about agents' attitudes in terms of their action selection based on their past experiences. Moreover, the tree is getting continuously updated based on both new received feedback and the outcome of chosen actions. This continuous update makes the approach adaptable to any change in the environment.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems.
The decision model provides web services with enough information which helps those services efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in their decision making and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. The proposed model can be seen as a recommneder system that suggests beneficial actions for both communities and single services. Communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact both their short-term and long-term goals; therefore, opting for the best decision available.

@@

\subsection{Contribution 3: Analyzing Coopetition Strategies of Services within Communities}

In previous contributions, our focus was on community formation and we emphasised on cooperative behavior of the web services as agents. In our next contribution\footnote{This contribution was published in \cite{DBLP:journals/eswa/AslBMKO14}} our focus is on the internal management affairs of communities of web services. Within communities, the web services, selfish and utility maximizers by nature, can follow two different strategies, namely cooperation and competition in order to increase their payoffs when they provide services to consumers \cite{VuFind-10008938119}. When competitive and cooperative behaviors and strategies are combined, autonomous services are said to be ``coopetitive''. In typical business settings, services are used to compete within communities as they provide the same functionalities and the number of users requests is finite. However, the same reason of providing similar functionalities can lead services to cooperate because they can replace each other in case of failure or unavailability, and services can do better in a coalition structure. Analyzing services competition and cooperation strategies within communities is still an open problem that motivates the research described in this section. we propose a mechanism within which service agents
in the community could choose either to compete for an announced task\footnote{Requests and tasks are used in this thesis interchangeably.}, or to cooperate with other competing services in the same community to accomplish some subtasks of the announced task. We equip intelligent web services to follow a reasoning technique to choose best interactive strategy (Coopetitive attitude, which is categorized to compete and cooperate). In the proposed system, We explore details behind the strategic decision making procedures and enable service agents to apply different techniques to constrain high efficiency and obtain the maximum utility. We investigate services' expected payoffs and the involved probabilities that are used to choose over the two interacting strategies.
